### 2021-07-06 rad-icfp-2021/complexity.tex
improving complexity section

As $\directD{\rho}{\Gamma}{Y}$ is defined by induction on programs, it suffices to show locally 
that $\directD{\rho}{\Gamma}{Y}$ verifies the cheap gradient principle. 
However, this is not quite the case with $\directD{\rho}{\Gamma}{Y}$ 
before the optimization (inlining and forward substitution) step from Section~\ref{sub:Partial evaluation and optimization}.

\begin{proposition}
    $\cost(\directD{\rho}{\Gamma}{Y}($!e!$))\leq$ $K\cost$(!e!) + O((0,0,0,n*|!e!|)) 
\end{proposition}

\begin{proof}
    TODO: routinely by induction and inspection. be specific with K? 
    One problem still is for map and reduce because they call twice the gradient of !e$_1$!.
\end{proof}

As we can see, the only problem is that too many variables are being moved, and this is
due to the continuation carrying and copying $n$ variables at each step.
This overhead is precisely what that our inlining and forward substitution removes, 
as was examplified in Section~\ref{sub:Partial evaluation and optimization}.

Combining the two previous points, we conclude that our reverse mode transformation satisfies the cheap gradient principle.

\begin{theorem}
    TODO: The reverse mode transformation satisfies the cheap gradient principle.
\end{theorem}

>>>>

As $\directD{\rho}{\Gamma}{Y}$ is defined by induction on programs, it suffices to show locally 
that $\directD{\rho}{\Gamma}{Y}$ verifies the cheap gradient principle.

There is a first restriction that prevents our transformation from satisfying the cheap gradient principle.
If we try to show that the cheap gradient principle holds by induction on the term, it fails on the !map2!.
When differentiating !map2!, there are two series of calls to (sub)gradients of !e$_1$!. 
The induction hypothesis is then too weak to conclude. 
The problem comes from the fact that !e$_1$! could itself use !map2!. 
So the constant $K$ can be independant of $n$ and of 
the size of the term if we allow it to be dependent on the level of nestings of !map2!.
A similar phenomenon happens with !reduce!.

Another problem is that the continuation part of $\directD{\rho}{\Gamma}{Y}$ adds $O(n)$ MOVES at each step.
This overhead is precisely what that our inlining and forward substitution removes, 
as was exemplified in Section~\ref{sub:Partial evaluation and optimization}.

Combining the two previous points, we conclude that our reverse mode transformation satisfies a version of the cheap gradient principle.

\begin{theorem}
    Given a term !$\Gamma \vdash$ e: $\reals$! in Source with at most $p$ nested !map2! or !reduce!. 
    Denote by $G$ the term $\grad_\Gamma$!e! after inlining and forward substitution.
    Then !$\cost$(G) $\leq$ $4*2^{p}*\cost$(e)!.
\end{theorem}

\begin{proof}[notes]
    The proof is by routine induction on !e!. 
    First, one computes the cost of $\directD{\rho}{\Gamma}{Y}$(!e!).
    It has $N$ too many MOVES which are removed by inlining and forward substitution using the invariant that at each step, 
    at most 2 !x!$_i$ in the continuation are not variables.
\end{proof}
